### 2012-11-28 Assignment6/paper/paper.tex
Consolidated data into 3 different tables, moved JSON and ER-diagram to be an appendix at the end of the paper, changed things in the tables so that they would fit into a single column.
If someone thinks more description is needed for a table try to put it into the
caption so that the tables themselves fit into a single column (it's a bit
cleaner this way since the numbers we have are not very big, it's just the
names of the columns, etc.)

Also, added the runtime for the assign task workload in couch.log

% Paper for the ChouchBase assignment.
\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{float}
%\usepackage{url}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

\begin{document}

\title{No Dogs On The Couch$_{Base}$}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Eriq Augustine, Ryan Hnarkis, Aldrin Montana, Ryan Verdon, Tyler Yero}
\\
\IEEEauthorblockA{Department of Computer Science\\
Cal Poly, San Luis Obispo\\
 \textsf{\{eaugusti, rhnaraki, amontana, rverdon, tyero\}@calpoly.edu}
}
}

\maketitle

\thispagestyle{empty}
\pagestyle{empty}

\section{Introduction}\label{sec:introduction}
Gene annotation is the process of associating metadata about a gene with the
contig, a dna sequence, on which the gene resides. This metadata is necessary for conducting
genomic research projects and analysis involving the contig's genomic sequence.
Currently, there is no specialty gene annotation software. Genome browsers
all maintain lots of genomic data and are used when manually annotating genes,
but do not provide a user-friendly interface. While genome browsers
will likely never be replaced (UCSC genome browser, etc. are well
established), it would be desirable to have a software system that better
accommodates the visual and informational needs of gene annotation.

\textit{CGAT} is a web-based gene annotation application designed with
usability, simplicity, and efficiency in mind. It is not a feature and data
rich genome browser like the \textit{UCSC genome brower}. Nor is \textit{CGAT}
designed to be a replacement for other genome browsers in any aspect other than
gene annotation. Other gene browsers attempt to accommodate many needs of the
biology community and so offer many features, and a cluttered, chaotic
interface. Gene annotation, while not a trivial task, can be well-serviced by a
simple data model. Ideally, by focusing on a simple data model, \textit{CGAT}
is able to provide a clean, streamlined experience. \textit{CGAT} aims to
provide users with a way to view gene annotations without extra, unnecessary
information obscuring their view while also embodying a Wikipedia-like emphasis
on collaboration and openness.

Anya Goodman, a professor in the biochemistry department at Cal Poly, San Luis
Obispo, offers a bioinformatics course that covers several aspects of gene
annotation. This project is spearheaded by Ryan Verdon to accommodate Dr.
Goodman's needs and ideas for ideal gene annotation software.

This paper primarily focuses on considerations for \textit{CGAT}s database
architecture. These considerations are made with regards to a MySQL cluster
versus a Couchbase cluster to determine which database architecture is most
efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
motivation to compare Couchbase and MySQL in this paper, while section
\ref{sec:design} talks about how we designed experimental workloads to use for
our comparison of Couchbase vs MySQL. In section \ref{sec:workload} we
describe the workloads and how they are relevant to \textit{CGAT}, some brief
implementation details are outlined in section \ref{sec:implementation}, we
evaluate our results in section \ref{sec:eval} and finally conclude in
section \ref{sec:conclusions}.

\section{Motivation}\label{sec:motive}
Traditional web applications are almost always backed by a relational database.
However, there has been a recent movement towards using NoSQL databases for
web application. Our main goal is to determine whether a NoSQL database, Couchbase,
can outperform a traditional relational database, MySQL, for \textit{CGAT}'s
common operations.

\section{Experiment Design}\label{sec:design}
To evaluate the performance we auto-generated around
400MB of data and inserted the same data into each cluster. Obviously using a
different data model for Couchbase to accommodate for the fact it is a NoSQL
key-value store instead of a relational database. We then created a set of six
experiments to run. Five came from common use cases found in the web
application. The final experiment was a contrived one designed to test the performance on each
cluster by repeatedly writing and reading from the same object. Our major
metrics are write performance, read performance, and time-delay between writes
and reads of the same data.

For our experiments we are using a total of five EC2 instances from the Amazon
Web Services cloud offering. For both of our systems four servers were used for the
database cluster while the last was used for running our workload harness. The MySQL
system is setup with one master server for writes and three slaves servers that
replicate all of the masters data. Unlike the Couchbase system where each
server handled a subset of the data.

Our major driving force of this experiment was to see how Couchbase compared to
MySQL for running a web application. To test which is better we created a
workload harness. The main tasks of the workload harness was to define how many
times a workload would run and which workloads to run. In our case workloads
define a certain task. For example, getting of the data necessary to display a
user's profile. In our system a workload knows how to do its job in both MySQL
and the Couchbase API. We leave describing each type workload we ran to the
next section.

As each workload is run the workload harness records important stats. Some of
the stats include how long the workload took to run, the min and max times of
an individual query or insertion, and the standard deviations of individual
runs.

\subsection{Data}
There exists four main pieces of data that are crucial to either
representation: contigs, annotations, users, and groups. Contigs are
what a biologist annotates to find genes. The main portion of a contig is a DNA
sequence. Annotations are the record or formal representation of a gene that
has been found in a specific contig. An annotation has several parts including
coordinates in the contig which specify individual parts of the gene, exons, and meta information
about the annotation like who uploaded it and when it was done, etc.

Groups are another important part of the usefulness of the system. Users of
like minds or research targets can join the same group to receive important information. Users is a
major part of the data and is one of the central pillars that connects the
data. Users upload contigs, create annotations, and join groups. Below we show
the different ways we store the data in SQL and JSON. We chose to represent all of our
data in Couchbase in JSON because it is a naturally serializable data format with
large library support. In addition, using JSON would allow use to share the same
schema with other popular document-based NoSQL databases, such as MongoDB.

\subsubsection{SQL}
In SQL the data is represented as nine tables. A diagram of the tables with
links to represent foreign key constraints is shown in Figure \ref{fig:SQLDiagram}.

\subsubsection{JSON}
For Couchbase we used four JSON objects which are shown in Figures \ref{fig:JSON-user}-\ref{fig:JSON-group}.
All of the information from the SQL tables is represented in these JSON
objects. We take advantage of nested objects to reduce as much subsequent
requerying of the database as possible.

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            {\textbf{Workload}} &
            {\textbf{MySQL}} &
            {\textbf{Couchbase}} \\
         \hline
            Task Assignment         & 39325189ms & 157431049ms \\
         \hline
            Profile View            & 2908257ms & 7008818ms \\
         \hline
            Annotation Publication  & 510837ms & 1830064ms \\
         \hline
            Group Modification      & 316673ms & 9086304ms \\
         \hline
            Contig Upload           & 119892ms & 20133ms \\
         \hline
            Read/Write              & 103808ms & 58581ms \\
         \hline
      \end{tabular}
      \caption{Comparison of runtimes in ms for workloads using MySQL and Couchbase.
               Each workload was run for 100000 iterations.}
      \label{tab:workload_perf}
\end{table}

\section{Workloads}\label{sec:workload}
The core unit of work that we used to compare Couchbase performance against
MySQL performance was called a \textit{workload}. Each workload was designed
either to be write-heavy, read-heavy, or balanced. The number of workloads used
during our comparison was determined by the number of members in our group (as
specified by Alex Dekhtyar). Unfortunately due to time constraints we were
unable to do multiple runs of any one workload, however, due to our environment
and machines, there should be minimal fluctuations in any given run time for a
workload.

In this section we further describe each task's purpose as well as what type of
workload it was (write-heavy, etc.). This is necessary to be able to properly
interpret (and agree or disagree with) the results described in Section
\ref{sec:eval}.

\subsection{Profile View}
The first thing that all users will do in \textit{CGAT} (ideally) is log in.
When a user logs into \textit{CGAT} the user will be directed to their profile
view which will contain all groups they are a member of, all completed
annotations they have submitted, partial annotations that have been saved, and
all tasks they have assigned. This is a use case that is likely to see many
additions in the future as it serves as the first and foremost view of the
system for the user. This means that any information that should be accessible
-- including when a forum system is integrated into \textit{CGAT} -- will find
a place on the user's profile.

This workload is clearly a read-heavy workload that entails no disk writes at
all. Important to note though, this use case may constantly be increasing
(should constantly be increasing for a successful system) and so the load for
this workload may increase quite a bit, and whenever new features are added
there is a potential for a spike in workload.

\subsection{Task Assignment}
In \textit{CGAT}, it must be possible to request a group of users to annotate
new or updated contigs. The frequency of this use case is dependent on the
adoption of \textit{CGAT} -- the more species \textit{CGAT} supports, the more
frequent this use case will be. More specifically, the relevant entities in the
\textit{CGAT} schema are \textit{users}, \textit{contigs}, and \textit{groups}.
Since a task embodies a request for a group $G$ to annotate a contig $C$, every
user $U$ in $G$ receives a notification including a description of $C$ and a
message requesting that it be annotated. A task can be assigned to any number
of groups. Since tasks are assigned to groups that have some vested interest in
the species or specific contig that has been updated or added, there may be
many groups that a task should be assigned to. Additionally, as more species
are supported by \textit{CGAT} or more users join \textit{CGAT}, more users
will receive notifications of tasks and so there are multiple ways in which
this use case must be scalable.

This workload is write-heavy -- adding tasks for each user in a group requires
adding many records to a relational table or modifying many documents/values
in a key/value store.

\subsection{Annotation Publication}
Annotations are the core focus of \textit{CGAT}. Every user is going to be
doing some annotation, and they should be doing many annotations. Especially
since \textit{CGAT} defines an annotation to be information about a gene
isoform, variation, on a contig, even annotating a single gene will yield to several
annotations in \textit{CGAT}. Additionally, when an annotation is submitted,
the annotation must be marked as complete, the user and annotation then are
assigned a certain amount of experience--the user gains experience towards
his total, the annotation gains experience and serves as a history of the
user's experience.

This is a mostly write-, somewhat read-heavy workload. It is write-heavy in the
sense that it must update a user and the annotation that the user completed.
However, it also has some reads in the sense that the contig, for which the
annotation is being submitted, is used to determine the amount of experience
that submitting the annotation yields for the user. However, given the
difference in data model between Couchbase and MySQL, the Couchbase
implementation is much more write-heavy than the SQL equivalent. This is
because the Couchbase data model includes lists of complete and incomplete
annotations for each user so that the annotation does not have to be fetched
for every use case involving users and annotations. Since this is the case,
more of the document undergoes modification for this use case.

\subsection{Group Membership Modification}
Group membership is likely the second most stable of the workloads used in our
comparisons of Couchbase vs MySQL next to contig uploads. This is ``stable''
because the majority of users will only join groups, and leaving groups will
likely be relatively rare. This use case will only ever be as bad as how many
users join at any one time--once a user registers for \textit{CGAT} and has
joined some groups the user is unlikely to join or leave groups and so this use
case poses almost no scalability problems. Although, despite this, there will
be spikes that coincide with academic schedules, as groups will be created and
populated for each new class that uses \textit{CGAT}.

This workload is the most balanced of the workloads described in this paper
(except maybe for ``Read/Write'' workload). Generally speaking, when a user joins a
group then there is only the addition of a record in MySQL or the modification
of the user and group document in Couchbase. And when a user leaves a group, then
subsequently the additional relational tuple or field in the user document is
``undone'' by removing the group membership tuple or deleting the field from
the user and group document.

\subsection{Contig Uploads}
Contig uploads is the most stable workload when the manual, or user, use case
is considered. Although, there will be large spikes of activity when/if the
reference genome for a species is updated--requiring \textit{CGAT} to update
any and all of the contigs relevant to the genome. However, reference genomes
are updated very infrequently and contigs will never be removed from
\textit{CGAT}.

This workload is write-heavy because it requires no reads--the use case is for
uploading data, not downloading data.

\subsection{Read/Write Performance}
This is a test workload to examine the performance of writing and reading from
the same object in succession. This is a workload contrived purely for better
comparison of Couchbase vs MySQL instead of workloads representative of what
\textit{CGAT} will rely on. This is a naive, balanced workload.

\section{Implementation}\label{sec:implementation}
Our experiment was implemented completely in Java. To do the MySQL portion of
the workload we used the MySQL JDBC driver. For the Couchbase portion we used
the official Couchbase Java library from their website.

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            \multicolumn{4}{|c|}{{\textbf{MySQL read/write performance}}} \\
         \hline
            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
         \hline
            Min (ms) & 0 & 0 & 0 \\
         \hline
            Max (ms) & 25 & 70 & 70 \\
         \hline
            Mean (ms) & 0.473 & 0.558 & 0.516 \\
         \hline
            Median (ms) & 0 & 1 & 0 \\
         \hline
            Standard Deviation (ms) & 0.685 & 0.859 & 0.778 \\
         \hline
      \end{tabular}
      \caption{Comparison of reads and writes to the same object in MySQL.}
      \label{tab:sql_perf}
\end{table}

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            \multicolumn{4}{|c|}{{\textbf{Couchbase read/write performance}}} \\
         \hline
            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
         \hline
            Min (ms) & 0 & 0 & 0 \\
         \hline
            Max (ms) & 71 & 48 & 71 \\
         \hline
            Mean (ms) & 0.562 & 0.017 & 0.290 \\
         \hline
            Median (ms) & 1 & 0 & 0 \\
         \hline
            Standard Deviation (ms) & 0.859 & 0.213 & 0.682 \\
         \hline
      \end{tabular}
      \caption{Comparison of reads and writes to the same object in couchbase}
      \label{tab:couch_perf}
\end{table}

\section{Evaluation}\label{sec:eval}
In this section we discuss the results from each workload in turn.

\subsection{Task Assignment}
From the results it is obvious that MySQL performs much better than Couchbase.
The result we have for Couchbase is actually a minimum runtime. It was running
so long we killed it to get more results. Even with this in mind, MySQL vastly
outperforms Couchbase. We believe this is because during the process of
assigning a task a couple values have to be read into memory before a finally
write can be done. Compared to MySQL where its a simple ad-hoc query to find
the necessary values prior to inserting the new task.

From our analysis, we believe if we had document or ad-hoc unstructured queries
in Couchbase we could vastly improve the performance.  When doing the Couchbase
queries we only need a small subset of the JSON. By only querying for the
necessary parts it could save vast amounts of time.

\subsection{Profile View}
Like the last workload, MySQL outperforms Couchbase. We believe that this is
because this workload is read-heavy so several full documents are retrieved per
run.  Again, this workload could be improved in Couchbase with the use of
unstructured queries.

\subsection{Annotation Publication}
Again, a similar pattern in the results as the last two workloads. This was
another workload that required some queries prior to inserting a new
annotation. We believe this is another case where unstructured queries could
vastly improve performance.

\subsection{Group Modification}
To our surprise this workload performed extremely bad on Couchbase. We were not
expecting this because this is a 50\% read and 50\% write workload. Having to
do lots of reads prior to inserting new group memberships really hurt
Couchbase. This is most definitely a case where having unstructured queries or
document support would greatly improve performance. The way we store group
memberships is as a sub-field in the user and group objects. By only queries
for those fields we could reduce the amount of data we return to the workload
harness by over 70\%.

\subsection{Contig Uploads}
The results from this test were no surprise. Couchbase demolished MySQL when
most of the workload was insertions. Even with contig objects containing
strings of over 50k characters Couchbase performed amazingly well. By not
having to lock key-value pairs Couchbase can get some awesome write
performance.

\subsection{Read/Write Performance}
The results from this test were not that surprising based on the results from
our other tests. It is obvious that Couchbase is much faster in terms of writes
plus Couchbase can leverage Memcache in this workload to speed up queries to
the database. For this workload we decided to gather a bunch more numbers to
see what the minimum, maximum, mean, median and standard deviation of the
individual reads and writes for each system. Below we list the six tables and
discuss what the results mean.

The three prior tables are all of our SQL results. The first table lists the
minimum, maximum, mean, median, and standard deviation of reads that occur
after writes. The second table lists the same metrics but for writes after
reads. Lastly, the third table lists the metrics for all of the queries. Below
we list the same three tables but this time show the Couchbase results.

By examining the read after writes results from the two systems we get the
results we expect based on our other tests. MySQL outperforms Couchbase when it
comes to reading after a write. The really interesting results come from the
the write after read tables where it is obvious that Couchbase demolished MySQL
by a lot. There is a huge difference in the means alone without even looking at
the other metrics. This again backs up our other results from the contig
workload where Couchbase outperformed MySQL in a mainly write task. Overall
Couchbase outperformed MySQL because of its blazing fast writes after reads.

\section{Conclusions}\label{sec:conclusions}
From our results MySQL appears to be the winner right now for the CGAT backend.
The MySQL workloads outperformed Couchbase the majority of the time. We believe
in a live scenario Couchbase would perform better than our tests mainly because
our tests could not take advantage of Memcache. When a key had to be looked up
there was almost no chance it was sitting in Memcache. We also believe that the
Couchbase results would greatly improve with the addition of an unstructured
query language. With the addition of this feature our workloads would be making
much smaller requests from the database.

% TODO: Uncomment if we use any refs
% \bibliographystyle{acm}
% \bibliography{refs}

\newpage
\appendices
\begin{figure*}
   \centering
   \includegraphics[height=90mm]{SQLDiagram.eps}
	\caption{The SQL diagram for CGAT.}
	\label{fig:SQLDiagram}
\end{figure*}

\begin{figure*}{t}
   \centering
   \subfloat[Contig JSON]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
   \subfloat[Annotation JSON]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
\end{figure*}

\begin{figure*}{t}
   \centering
   \subfloat[User JSON]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
   \subfloat[Group JSON]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
\end{figure*}

%\begin{figure*}
%  \includegraphics[height=90mm]{user.eps}
%	\caption{The JSON for a user.}
%	\label{fig:JSON-user}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=25mm]{group.eps}
%	\caption{The JSON for a group.}
%	\label{fig:JSON-group}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=80mm]{contig.eps}
%  \caption{The JSON for a contig.}
%  \label{fig:JSON-contig}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=90mm]{annotation.eps}
%	\caption{The JSON for an annotation.}
%	\label{fig:JSON-annotation}
%\end{figure*}

\end{document}